{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-07-15 thesis/tm.tex
synhi

\begin{figure}[t]
	\begin{center}
		% TODO synhi
		\begin{tabular}{l}
		\texttt{bool hle\_spinlock\_isfree(spinlock\_t* lock) \{} \\
		\texttt{~~~~// XXX: should be "return lock->v == 0;"} \\
		\texttt{~~~~return \_\_sync\_bool\_compare\_and\_swap(\&lock->v, 0, 0);} \\
		\texttt{\}} \\
		\texttt{void rtm\_spinlock\_acquire(spinlock\_t* lock) \{} \\
		\texttt{~~~~if ((tm\_status = \_xbegin()) == \_XBEGIN\_STARTED) \{} \\
		\texttt{~~~~~~~~if (hle\_spinlock\_isfree(lock)) return;} \\
		\texttt{~~~~~~~~\_xabort(0xff);} \\
		\texttt{~~~~\} else \{} \\
			\texttt{~~~~~~~~// ... retrying \&c abbreviated for brevity ...} \\
		\texttt{~~~~~~~~hle\_spinlock\_acquire(lock);} \\
		\texttt{~~~~\}} \\
		\texttt{\}} \\
		\texttt{void rtm\_spinlock\_release(spinlock\_t* lock)\{} \\
		\texttt{~~~~if (hle\_spinlock\_isfree(lock)) \{} \\
		\texttt{~~~~~~~~\_xend();} \\
		\texttt{~~~~\} else \{} \\
		\texttt{~~~~~~~~hle\_spinlock\_release(lock);} \\
		\texttt{~~~~\}} \\
		\texttt{\}} \\
		\end{tabular}
	\end{center}
	% TODO describe
	\caption{Code from {\tt spinlock-rtm.c}, modified only to remove unrelated logic for brevity,
		showing the performance bug Landslide found in {\tt lock\_fast}({\tt spinlock}).
		The {\tt isfree()} routine uses an atomic read-and-write operation where just a read would suffice,
		which leads to superfluous memory conflicts in the transactional path
		(seen at both of its callsites below).
	}
	\label{fig:spinlockbug}
\end{figure}

@@

\begin{figure}[h]
	% TODO: syn hi
	\begin{center}
	\begin{tabular}{p{0.47\textwidth}p{0.5\textwidth}}
		\begin{tabular}{l}
			\texttt{static int num\_in\_section = 0;} \\
			\texttt{for (int i = 0; i < NITERS; i++) \{} \\
			\texttt{~~~~rtm\_spinlock\_acquire(\&lock);} \\
			\texttt{~~~~num\_in\_section++;} \\
			\texttt{~~~~if (!\_xtest())} \\
			\texttt{~~~~~~~~thr\_yield(-1);} \\
			\texttt{~~~~assert(num\_in\_section == 1); } \\
			\texttt{~~~~num\_in\_section--;} \\
			\texttt{~~~~rtm\_spinlock\_release(\&lock);} \\
			\texttt{\}} \\
		\end{tabular}
		&
		\begin{tabular}{ll}
			\texttt{for (int i = 0; i < NITERS; i++) \{} \\
			\texttt{~~~~rtm\_spinlock\_acquire(\&lock);} \\
			\texttt{~~~~assert(\_xtest());} \\
			\texttt{~~~~rtm\_spinlock\_release(\&lock);} \\
			\texttt{\}} \\
		\end{tabular}
		\\
		\\
		(a) {\tt lock}(), tests mutual exclusion.
		&
		(b) {\tt lock\_fast}(), tests for no spurious aborts.
	\end{tabular}
	\end{center}
	\caption{Abstraction reduction test cases.}
	\label{fig:htm-lock}
\end{figure}

Table~\ref{tab:verifs2} shows the new resulting levels of verification
Landslide reached before the same 10-hour timeout.
Provided that one trusts the {\tt lock} tests correctly check the desired properties,
and that open-coding hadn't introduced any new bugs (such as Figure~\ref{fig:avlbug}'s),
the benefit is clear:
for example,
summing the testing times of {\tt lock}({\tt mutex})$(K,N)$ and {\tt avl\_mutex}$(K,N)$
produces the same verification as {\tt avl\_fixed}$(K,N)$ far more cheaply.
Furthermore, {\tt lock}'s verification can be reused,
whereas {\tt avl\_fixed} and {\tt map\_basic} effectively duplicated the mutex verification between them.
Note also the impact that fixing Figure~\ref{fig:spinlockbug}'s performance bug
(changing a read+write to a read only)
had on even the correctness tests:
{\tt lock}({\tt spin\_fixed})'s state spaces were reduced by more than half compared to {\tt lock}({\tt spinlock}).

>>>>

\begin{figure}[t]
	\begin{center}
		% TODO synhi
		\begin{tabular}{l}
		\texttt{\ctype{bool} \call{hle\_spinlock\_isfree}(\ctype{spinlock\_t} *lock) \{} \\
		\texttt{~~~~\ccomment{// XXX: should be "return lock->v == 0;"}} \\
		\texttt{~~~~\flow{return} \call{\_\_sync\_bool\_compare\_and\_swap}(\&lock->v, \const{0}, \const{0});} \\
		\texttt{\}} \\
		\texttt{\ctype{void} \call{rtm\_spinlock\_acquire}(\ctype{spinlock\_t} *lock) \{} \\
		\texttt{~~~~\flow{if} ((tm\_status = \call{\_xbegin}()) == \const{\_XBEGIN\_STARTED}) \{} \\
		\texttt{~~~~~~~~\flow{if} (\call{hle\_spinlock\_isfree}(lock)) \flow{return};} \\
		\texttt{~~~~~~~~\call{\_xabort}(\const{0xff});} \\
		\texttt{~~~~\} \flow{else} \{} \\
		\texttt{~~~~~~~~\ccomment{// ... retrying \&c abbreviated for brevity ...}} \\
		\texttt{~~~~~~~~\call{hle\_spinlock\_acquire}(lock);} \\
		\texttt{~~~~\}} \\
		\texttt{\}} \\
		\texttt{\ctype{void} \call{rtm\_spinlock\_release}(\ctype{spinlock\_t} *lock)\{} \\
		\texttt{~~~~\flow{if} (\call{hle\_spinlock\_isfree}(lock)) \{} \\
		\texttt{~~~~~~~~\call{\_xend}();} \\
		\texttt{~~~~\} \flow{else} \{} \\
		\texttt{~~~~~~~~\call{hle\_spinlock\_release}(lock);} \\
		\texttt{~~~~\}} \\
		\texttt{\}} \\
		\end{tabular}
	\end{center}
	% TODO describe
	\caption{Code from {\tt spinlock-rtm.c}, modified only to remove unrelated logic for brevity,
		showing the performance bug Landslide found in {\tt lock\_fast}({\tt spinlock}).
		The {\tt isfree()} routine uses an atomic read-and-write operation where just a read would suffice,
		which leads to superfluous memory conflicts in the transactional path
		(seen at both of its callsites below).
	}
	\label{fig:spinlockbug}
\end{figure}

@@

\begin{figure}[h]
	% TODO: syn hi
	\begin{center}
	\begin{tabular}{p{0.47\textwidth}p{0.5\textwidth}}
		\begin{tabular}{l}
			\texttt{\ctype{static int} num\_in\_section = 0;} \\
			\texttt{\flow{for} (\ctype{int} i = \const{0}; i < \const{NITERS}; i++) \{} \\
			\texttt{~~~~\call{rtm\_spinlock\_acquire}(\&lock);} \\
			\texttt{~~~~num\_in\_section++;} \\
			\texttt{~~~~\flow{if} (!\call{\_xtest()})} \\
			\texttt{~~~~~~~~\call{thr\_yield}(\const{-1});} \\
			\texttt{~~~~\flow{assert}(num\_in\_section == \const{1}); } \\
			\texttt{~~~~num\_in\_section--;} \\
			\texttt{~~~~\call{rtm\_spinlock\_release}(\&lock);} \\
			\texttt{\}} \\
		\end{tabular}
		&
		\begin{tabular}{ll}
			\texttt{\flow{for} (\ctype{int} i = \const{0}; i < \const{NITERS}; i++) \{} \\
			\texttt{~~~~\call{rtm\_spinlock\_acquire}(\&lock);} \\
			\texttt{~~~~\flow{assert}(\_xtest());} \\
			\texttt{~~~~\call{rtm\_spinlock\_release}(\&lock);} \\
			\texttt{\}} \\
		\end{tabular}
		\\
		\\
		(a) {\tt lock}(), tests mutual exclusion.
		&
		(b) {\tt lock\_fast}(), tests for no spurious aborts.
	\end{tabular}
	\end{center}
	\caption{Abstraction reduction test cases.}
	\label{fig:htm-lock}
\end{figure}

Table~\ref{tab:verifs2} shows the new resulting levels of verification
Landslide reached before the same 10-hour timeout.
Provided that one trusts the {\tt lock} tests correctly check the desired properties,
and that open-coding hadn't introduced any new bugs (such as Figure~\ref{fig:avlbug}'s),
the benefit is clear:
the data structure tests suddenly reach unprecedented thread/iteration counts with ease,
their state space growth now defined only by their internal conflicts from tree rebalancing, map resizing, and so on.
In total,
summing the testing times of {\tt lock}({\tt mutex})$(K,N)$ and {\tt avl\_mutex}$(K,N)$
produces the same verification as {\tt avl\_fixed}$(K,N)$ far more cheaply.
Furthermore, {\tt lock}'s verification can be reused,
whereas {\tt avl\_fixed} and {\tt map\_basic} effectively duplicated the mutex verification between them.
Note also the impact that fixing Figure~\ref{fig:spinlockbug}'s performance bug
(changing a read+write to a read only)
had on even the correctness tests:
{\tt lock}({\tt spin\_fixed})'s state spaces were reduced by more than half compared to {\tt lock}({\tt spinlock}).
